---
title: "DOM EEM processing"
subtitle: "Template for Analysis"
author: "Emily Taylor"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{placeins}
  - \usepackage{subfig}
  - \usepackage{booktabs}
  - extra_dependencies: subfig
output: 
  html_notebook: 
    toc_depth: 5
    toc: true
    toc_float: true
    number_sections: yes
    df_print: paged
    code_folding: hide
  pdf_document:
    fig_crop: yes
    toc: true
    number_sections: true
geometry: margin = 0.75in
editor_options:
  markdown:
    wrap: 80
  chunk_output_type: console
---

More details on the protocols followed for building this PARAFAC model can be found in the [StaRdom package vingette](https://cran.r-project.org/web/packages/staRdom/vignettes/PARAFAC_analysis_of_EEM.html)


#testing - 
```{r}
#Set your main Aqualog folder directory where EVERYTHING lives (projects, scripts, raw data)
setwd("01_Raw_data/Aqualog_processing")

#name of the project you are working in today
project_name <- "04162024"

#If you diluted, name the dilution csv that will be created later
dilution_sheet_name <- paste("metatable_dilution_",project_name,".csv",sep = "")

#Pick a sample that you want to designate as your "check" sample to look at . if we were to revie
review_sample <- c("5_11012023")
#review_sample <- "13"

#Today's date so that output files can have the date in it
output_date <- "04162024"

#create output folder if you dont have one
#dir.create(paste("./projects",project_name,"output",sep = "/"))
```


# Getting started: libraries, loading custom functions, creating dilution sheet

1. Libraries, working directory
```{r library, include = FALSE, message = FALSE}
librarian::shelf(
  tidyr,
  plyr,
  dplyr,
  staRdom,
  parallel, 
  doParallel,
  foreach,
  rlist,
  plot3D,
  data.table,
  knitr,
  quiet = TRUE)

options(scipen = 999)

library(tcltk)
```

2. load functions
```{r functions, include = FALSE}

source("./scripts/functions/format_aqua_data.R")

source("./scripts/functions/plotting_parameters_and_functions.R")

source("./scripts/functions/spectral_corrections.R")

source("./scripts/functions/grid_size_adjustment.R")

source("./scripts/functions/eem_remove_scattering_nonsym.R")

#Emily's updated FI index and associated functions
msg_warning_wavelength <- function() {
  msg <- "This metric uses either excitation or emission wavelengths that were not present in the data. Data has been interpolated to fit the requested wavelengths."
  return(msg)
}

is_eemlist <- function(eem) {
  ifelse(class(eem) == "eemlist", TRUE, FALSE)
}

is_eem <- function(eem) {
  ifelse(class(eem) == "eem", TRUE, FALSE)
}

eem_fluor_index <- function(eem, verbose = TRUE) {
  library(staRdom)
  
  stopifnot(is_eemlist(eem) | is_eem(eem))
  
  ## It is a list of eems, then call lapply
  if (is_eemlist(eem)) {
    res <- lapply(eem, eem_fluor_index, verbose = verbose)
    res <- dplyr::bind_rows(res)
    
    return(res)
  }
  
  if (!all(370 %in% eem$ex & c(470, 520) %in% eem$em) & verbose) {
    warning(msg_warning_wavelength(), call. = FALSE)
  }
  
  fluo_470 <- pracma::interp2(eem$ex, eem$em, eem$x, 370, 470)
  fluo_520 <- pracma::interp2(eem$ex, eem$em, eem$x, 370, 520)
  
  fi_cor <- fluo_470 / fluo_520
  
  return(data.frame(sample = eem$sample, fi_cor = fi_cor, stringsAsFactors = FALSE))
}
```

3. creat abs and eem folders if you dont already have them within the project folder
```{r}
#dir.create(paste("./projects",project_name,"eem",sep = "/"))
#dir.create(paste("./projects",project_name,"abs",sep = "/"))

```

4. format data
```{r}
convert_aqualog()
#1. select the project folder
#2. select individual analysis files within the raw-data folder, one at a time
```


#1. Load in EEMs and ABS Data

We will also generate and load in a metatable with dilution factors
which may or may not have been used during sample analysis. The current
function does not write the file path for absorbance data correctly, so
that function is also re-written prior to generating the metatable
template.
```{r }

#What is this for??
cores <- parallel::detectCores(logical = FALSE)


#select project folder directory where abs and eem folders are stored
data_folder <- choose_directory()

eem_folder <- paste(data_folder,"/eem",sep="") # selects eem folder
eem_list <- eem_read(eem_folder, import_function = eem_csv, recursive = T) # list of all eem files

absorbance_path <- paste(data_folder,"/abs",sep="") # select abs folder
absorbance <- absorbance_read(absorbance_path, recursive = T, cores = cores) # load csv or txt tables in folder

```


Reviewing the designated review sample. 
If you want to look at the scans of every sample, run the code that has eem_list inside the plotting function
Reviewing every single sample will take more time to produce figures but may be worth it to make sure samples are 
```{r}

#selecting eem files for only those you want to review
 eem_list_forReview <-
  eem_extract(
    eem_list,
   # c("B1S1S1"), enter others here
   # "blank", use this to look at all the blanks
   review_sample,
    keep = TRUE,
    ignore_case = TRUE)

#plot the review sample 
eem_overview_plot(eem_list_forReview)   

```

# Create, fill in, and import DILUTION metatable
the dilution table has a column "dilution" with 1 (no dilution). change this value if you diluted, leave 1 if not. leave 1 for the "_blank" sample
```{r meta-import}

#Step 1: create dilution sheet
eem_metatemplate(eem_list, absorbance) %>%
  mutate(dilution = "1") %>% 
   write.csv(file = paste("./projects",dilution_sheet_name,sep = "/"), row.names = FALSE)

`#Step 2: go open that dilution sheet and fill it in manually!!!!

library(tidyverse)
dil_table <- read_csv("projects/metatable_dilution_04162024.csv")
dil_table$site_date<-str_split_fixed(dil_table$sample, "[A-Z]+", 4)[,4]
dil_table$site<-str_split_fixed(dil_table$site_date, "_", 2)[,1]
dil_table$Date<-str_split_fixed(dil_table$site_date, "_", 2)[,2]
dil_table$Date<-mdy(dil_table$Date)
dil_table<-dil_table[,-4]

file.names <- list.files(path="projects/Bradford Streams/sample logs", pattern=".csv", full.names=TRUE)
logs_all<-data.frame()
for(i in file.names){
  log<-read_csv(i)
  log<-log[,c(2,3,4)]
  log<-rename(log, 'site'='Sample ID', 'Date'='Sampled')
  logs_all<-rbind(logs_all, log)}
logs_all$Date<-mdy(logs_all$Date)

dilution<-left_join(dil_table, logs_all, by=c('site','Date'))
dilution<-dilution[,-c(4,5,6)]
write.csv(dilution,"projects/Bradford Streams/metatable_dilution_04162024.csv")

#Step 3: read in edited dilution sheet
meta <-
  read.table(paste("./projects/Bradford Streams",dilution_sheet_name,sep = "/"),
    header = TRUE,
    sep = ",",
    dec = ".",
    row.names = 1)

```

#1. Remove Questionable Samples

Remove any samples with known issues such as poor data collection,
undiluted and oversaturated CCD, or just visually funky samples.
```{r remove-problem-samples}
# eem_list <-
#   eem_extract(
#     eem_list,
#     c("19MAY31",
#       "19MAY24",
#       "B2S20S20"),
#     ignore_case = TRUE)

# remove samples that exceeded CCD saturation. Some samples will have a diluted sample but a lot of the early ones do not
eem_list <-
  eem_extract(
    eem_list,
    c(
      "B1S7_HAT_2019Feb20",
      "B1S10_HOGUP_2019Aug14",
      "B1S11_HAT_2019Apr03" 
    ),
    ignore_case = TRUE
  )

# you can also fix sample names
eem_list <- 
  eem_name_replace(eem_list,
                   c("^B1S1_SWBUP_2019Jun05$"),
                   c("^B1S1_SWBUP_2019Jun18$"))

```

#1. Apply Corrections to ABS Data

Absorbance baseline correction corrects the instrumental baseline drift in absorbance data by subtracting the mean of the absorbance at high wavelengths (Li and Hur 2017). Default range is `c(680, 700)` but can be changed (i.e., `c(650, 800)`)

```{r load-abs-data}
absorbance <- abs_blcor(absorbance, wlrange = c(680, 700))
```

#1. Fix Emission Step Size

This adjusts the increments between emission wavelengths so that they are uniform across the spectra. The m-correct and x-correct files do not need to be applied as that has already been done by the software.

Try to match your wavelength range and step size increments to the range
and step sizes you collected the data at.

```{r fix-grid-size, message = FALSE, warning = FALSE, fig.show = 'hold'}

#NOT WORKING - this function is not changing all objects of a sample's list (needs to change x, em, and ex, only changes x)
#eem_list <-
#  adjust_fluorescence(eem_list, emission_range = c(246, 827), excitation_range = c(239, 800), emission_step = 2.5,excitation_step = 3)

#WORKING - for now I am using the old approach here, but removing the emcor part and using Emily's ranges
eem_list <- eem_range(eem_list,ex = c(239,800), em = c(246,800))

#double checking the min and max here
min(eem_list[[1]][["ex"]]) #239
max(eem_list[[1]][["ex"]]) #800

#selecting eem files for only those you want to review
 eem_list_forReview <-
  eem_extract(
    eem_list,
   review_sample,
    keep = TRUE,
    ignore_case = TRUE)

```

#1. Blank Subtraction

You may need to adjust the spectra range because blanks must be the same length as the samples they will be extracted from
```{r cut-range, echo = TRUE}
eem_list <-
  eem_range(eem_list, 
            ex = c(245, 600), 
            em = c(245, 650)
            )
```

```{r blank-subtraction}
# remove the blank scatter from sample scatter. if you have more than one blank, they will be averaged before removal. 
eem_list <- eem_remove_blank(eem_list)

# if removing a blank introduces some strongly negative values, a) double check your blank is good, if it is then b) try replacing negative values with 0
eem_list <- eem_interp(eem_list, type = 0, nonneg = TRUE)

 eem_list_forReview <-
  eem_extract(
    eem_list,
   review_sample,
    keep = TRUE,
    ignore_case = TRUE)
# View EEMs plots
eem_overview_plot(eem_list_forReview, spp = 9, contour = TRUE)

```

#1. Inner filter effect (IFE) Correction

```{r ife-correction, message = FALSE}
eem_list <-
  eem_ife_correction(
    data = eem_list,
    abs_data = absorbance,
    cuvl = 1,
    unit = "absorbance"
  )
#the warning you get for the "_blank" samples is normal. ignore



#review data for 1 site
eem_list_forReview <-
  eem_extract(
    eem_list,
    review_sample,
    keep = TRUE,
    ignore_case = TRUE)

# View EEMs plots
eem_overview_plot(eem_list_forReview, spp = 9, contour = TRUE)

```

#1. Raman Normalization

```{r raman-norm}
eem_list <- eem_raman_normalisation2(eem_list, blank = "blank")

 eem_list_forReview <-
  eem_extract(
    eem_list,
    review_sample,
    keep = TRUE,
    ignore_case = TRUE)

# View EEMs plots
eem_overview_plot(eem_list_forReview, spp = 9, contour = TRUE)

```

#1. Remove Blanks From Sample Set
```{r blank-sample-removal}
eem_list <-
  eem_extract(eem_list,
              c("nano", "miliq", "milliq", "mq", "blank"),
              ignore_case = TRUE)

absorbance <-
  dplyr::select(absorbance,
                -dplyr::matches("nano|miliq|milliq|mq|blank", ignore.case = TRUE))
```

#1. Removal Raman and Rayleigh Scatter

```         
- The function order is "raman1", "raman2", "rayleigh1" and "rayleigh2" 
- The figure order is raman 2, then rayleigh2, then raman 1, then rayleigh 1
```

#### 1. Check Raman and Rayleigh Bands

This is a side view of the emission fluorescence at excitation = 350. This can be used to assess how wide the Raman and Rayleigh scattering bands are and whether the chosen scattering removal band covers their full width
-TBD Audrey has to check this section
```{r check-em-at-350, fig.show = 'hold'}

# eem_bands(eem_list_done2)

```

Remove Rayleigh Band 2 
```{r remove-raman-and-rayleigh-scatter}
eem_list_rem <-
  eem_rem_scat(
    eem_list,
    remove_scatter = c(FALSE, FALSE, FALSE, TRUE),
    remove_scatter_width = c(5, 5, 0, 20),
    interpolation = TRUE
  )

 eem_list_rem_forReview <-
  eem_extract(
    eem_list_rem,
    review_sample,
    keep = TRUE,
    ignore_case = TRUE)
 
# Check Band Removal Widths
eem_overview_plot(eem_list_rem_forReview, spp = 9, contour = TRUE)

```

Remove Rayleigh Band 1 
```{r remove-scatter-asymmetrically}
eem_list_rem <-
  eem_remove_scattering_nonsym(
    eem_list_rem,
    type = "rayleigh",
    order = 1,
    width_above = 10, #change this and the line below if you want a thinner/thicker chunk removed
    width_under = 50
  )

 eem_list_rem_forReview <-
  eem_extract(
    eem_list_rem,
    review_sample,
    keep = TRUE,
    ignore_case = TRUE)
 
# Check Band Removal Widths
eem_overview_plot(eem_list_rem_forReview, spp = 9, contour = T)

```

Interpolate Scattering
```{r interp-EEMs, warning = FALSE}
# Interpolate EEMs for removed scatter bands
eem_list_done <- 
  eem_interp(
    eem_list_rem,
    cores = cores,
    type = 0,
    verbose = TRUE,
    nonneg = TRUE
  )


eem_list_done_forReview <-
  eem_extract(
    eem_list_done,
    review_sample,
    keep = TRUE,
    ignore_case = TRUE)
 
# View interpolated EEMs
eem_overview_plot(eem_list_done_forReview, spp = 9, contour = T)

```

#1. Applying dilution data
```{r}

dil_data <- meta["dilution"]

eem_list_done <- eem_dilution(eem_list_done,dil_data)

#eem_list_dil <- eem_list_rem_int

#applying dilution to absorbance
abs_dil <-
  eem_absdil(absorbance,
             eem_list = eem_list_done,
             dilution = dil_data,
             cor_data = NULL,
             auto = TRUE,
             verbose = FALSE)
```


#Calculating indices
```{r}
#abs
slope_parms <-
  abs_parms(abs_dil, #absorbance, [INPUT EITHER THE DILUTED OR NON DILUTED DATA - HERE THE DILUTED ONE IS SELECTED]
            cuvle = 1,
            unit = "absorbance",
            add_as = c(250, 272, 340, 350, 365, 400, 405),
            limits = list(c(325, 375), c(275, 295), c(350, 400), c(300, 700)),
            l_ref = list(325, 275, 350, 300),
            cores = cores
  )

slope_parms <- slope_parms %>%
  mutate(E3_E4 = a340 / a254) # has been used to predict DOC; range: 0.20 - 0.38 (Tipping et al., 2009)

write.csv(slope_parms, row.names = FALSE, 
          file = paste("./projects/Bradford Streams/output/indices_abs_",output_date,".csv",sep = ""))

#eem
eem_index <-
  cbind.data.frame(
    eem_biological_index(eem_list_done),
    eem_coble_peaks(eem_list_done),
    eem_fluorescence_index(eem_list_done), # old method
    eem_fluor_index(eem_list_done), # new method
    eem_humification_index(eem_list_done, scale = FALSE))

write.csv(eem_index, row.names = FALSE, 
          file = paste("./projects/Bradford Streams/output/","indices_eem_",output_date,".csv",sep = ""))

```





#left join vial IDs
```{r clean-up-and-save}
eem
library(stringr)
B1S1S3_12072023 
eem$site_date<-str_split_fixed(eem$sample...1, "[A-Z]+", 4)[,4]
eem$site<-str_split_fixed(eem$site_date, "_", 2)[,1]
eem$Date<-str_split_fixed(eem$site_date, "_", 2)[,2]
```
